Estimating the dynamics of SPARQL query results

### 2019-04-11 kgswc19.tex
an update

\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
\hypersetup{
    colorlinks, linkcolor={dark-red},
    citecolor={dark-green}, urlcolor={dark-blue},
    pdftitle={Estimating the dynamics of SPARQL query results},    % title
    pdfauthor={Alberto Moya, Aidan Hogan},     % author
    pdfsubject={KGSWC 2019},   % subject of the document
    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
}

@@

\begin{document}
%
\title{Estimating the dynamics of SPARQL query results\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{IMFD; DCC, University of Chile\\
\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We propose a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-measure for the predictions made; for reference, we include a baseline that randomly guesses yes/no. With the highest result being an $F_1$-measure of 0.706, the results show that predicting changes in query results is indeed a challenging problem; on the other hand, all configurations clearly outperform the random baseline. Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available. However, the \textsc{qp} features -- which do not assume the availability of such information -- are often competitive: more specifically, the best result for \textsc{r} is 0.706 with Linear SVM, while the best results for \textsc{qp} is 0.651 with Decision Trees.

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} and \textsc{tensorflow} libraries. The latter allows us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques. \ah{Could do with a comparison here; the 0.43 result is quite abstract.}

@@

@@

	\end{axis}
	\end{tikzpicture}
	\caption{Evolution of total triples in Wikidata (left) and number of triples added to and removed from the $n$\textsuperscript{th} version for the $n+1$\textsuperscript{th} version (right) \label{fig:result}}
\end{figure}

@@

In this paper we propose methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future. More specifically, we propose a framework based on classifiers and regression that accept features extracted from the query, from past versions of the data, and/or from the combination of both. Considering this framework, we propose that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.

@@

Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for predictions in the OSC and TTL tasks. For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions. Our results show that although features based on historical changes to query results perform best ($F_1 = 0.706$), considering static query features and predicate dynamics alone is competitive ($F_1 =0.651$) when historical results are not available or are prohibitively costly to compute.

>>>>

\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
\hypersetup{
    colorlinks, linkcolor={dark-red},
    citecolor={dark-green}, urlcolor={dark-blue},
    pdftitle={Estimating the dynamics of SPARQL query results},    % title
%    pdfauthor={Alberto Moya, Aidan Hogan},     % author
    pdfsubject={ISWC 2019},   % subject of the document
    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
}

@@

\begin{document}
%
\title{Estimating the dynamics of SPARQL query results
%\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
%Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
%\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{IMFD; DCC, University of Chile\\
%\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We propose a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$ -measure for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9 , 17); for reference, we include a baseline that randomly guesses yes / no. 
We can notice the best results were obtained using the characteristics of R reaching values ​​of 0.83 of f1. The results show that predicting changes in query results is indeed a challenging problem. Looking in further detail, as can be expected, prediction performs best when the \ textsc {r} feature - with knowledge of changes in the historical value of results of the input query - is available. And, the features from sets \textsc {q} and  \textsc {p} - which do not assume the availability of such information - obtain less F1-score: more specifically, the best result using \ textsc {pr} is 0.83 with Linear SVM, window size 17, while the best results without \ textsc {r} is 0.60 with Linear SVM, window size 5.

Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering r, being better as the window increases. However, these results are not affected when we consider only predicates in q or p. Also, we show, on the other hand, that all configurations outperform the random baseline.

@@

	\caption{$F_1$-measure for tested classifiers considering different sets of features (Q, P, R) and varying the the window sizes (2,5,9,17)}
	\label{tab:classifier}

@@

	\end{axis}
	\end{tikzpicture}
	\caption{Best results for each feature sets in term of $F_1$-measure (left), Precision score (center) and Recall (right) with different windows sizes 
		\label{fig:result}}
\end{figure}

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn}. The latter allows us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques.

@@

In this paper we evaluate methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future. More specifically, we evaluate a framework based on classifiers and regression that accept features extracted from the query, from past versions of the data, and/or from the combination of both. Considering this framework, we show that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.

@@

Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for predictions in the OSC and TTL tasks. For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions. Our results show that the features based on historical changes to query results perform best ($F_1 = 0.83$), considering static query features and predicate dynamics alone is less competitive ($F_1 =0.60$) when historical results are not available or are prohibitively costly to compute.
